---
title: "International Merchandise Trade Statistics Highlights"
author: Christophe Bontemps (SIAP)^[Based on data and information from Vanuatu National
  Statistical Office (VNSO) ]
subtitle: '*A toy model!*'
output:
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    keep_md: yes
    code_folding: show
    fig_width: 6.5
    fig_height: 4
  pdf_document:
    df_print: kable
    toc: no
    keep_tex: yes
    fig_width: 6.5
    fig_height: 4
---


```{r setup, include=FALSE}
# Set echo=FALSE as default for code cells
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      dpi = 600)
# My colors:
SIAP.color <- "#0385a8"

```


```{r libraries, include = FALSE}
#### Preparation ####

# Load the required libraries
library(dplyr) # Manipulating data
library(stringr) # common string operations
library(tidyverse) # Manipulating data

# For tables 
library(kableExtra)
library(modelsummary)
library(gt) 
library(gtsummary)
library(tinytable)


 # to get tinytable library(remotes) install_github("vincentarelbundock/tinytable")
```


```{r data, include = FALSE}
# # Note where VNSO code/data is on current computer
# repository <- file.path(dirname(rstudioapi::getSourceEditorContext()$path), "Data")
# setwd(repository) # Required for file.choose() function

# Read in the raw trade data from secure folder of the repository 
tradeStatsFile <- "VBoS_Dummy-Data_16.02.23.csv"
tradeStats <- read.csv(paste0("Data/", tradeStatsFile), header=TRUE, na.strings=c("","NA", "NULL", "null")) #replace blank cells with missing values-NA
```

# Data used 
We use some anaonyized data from Vanuatu National Statistical Office (VNSO) trade data.  and specifically used the file named  *`r tradeStatsFile`*, with  **`r nrow(tradeStats)`** observations on several goods exported.

```{r preparation, include = FALSE}
#### Clean and process the latest month's data ####

# Initial re-formatting of the data

# Remove duplicated rows from the trade statistics data
duplicatedRows <- duplicated(tradeStats) 
tradeStatsNoDup <- tradeStats[duplicatedRows == FALSE, ]

# Convert the statistical value to numeric - note numbers formatted with commas in them and these need removed
tradeStatsNoDup$Stat..Value <- as.numeric(gsub(",", "", tradeStatsNoDup$Stat..Value))

# Convert excel figures to dates (choose)
tradeStatsNoDup$Reg..Date <- as.Date(tradeStatsNoDup$Reg..Date, format = "%d/%m/%Y")

# Create subset for Export and Import commodities 
tradeStatsSubset <- tradeStatsNoDup[tradeStatsNoDup$Type %in% c("EX / 1","EX / 3", "IM / 4", "IM / 7"), ]
tradeStatsCommodities <- tradeStatsSubset[tradeStatsSubset$CP4 %in% c(1000, 4000, 4071, 7100), ]
```


```{r NewVariables, include = FALSE}
#### Extract the Year, Month, and Day as separate columns ####

# Create new columns for dates
tradeStatsCommodities$Year <- format(tradeStatsCommodities$Reg..Date, format= "%Y")
tradeStatsCommodities$Month <- format(tradeStatsCommodities$Reg..Date, format= "%B")
tradeStatsCommodities$Day <- format(tradeStatsCommodities$Reg..Date, format= "%d")

#### Finish Processing ####

# Make copy of latest month's processed data
processedTradeStats <- tradeStatsCommodities

# Create csv of last months processed data
write.csv(processedTradeStats, "Output/OUT_PROC_ASY_ProcessedRawData_Date.csv")

```

```{r, eval=TRUE}
head(processedTradeStats)
```


### Addition to the original file

The dates and year are missing, so we'll generate the year to have some variabilty
```{r}
 processedTradeStats  <- processedTradeStats %>%
  mutate (Year = as.integer(2000 + round(10*runif(nrow(.)), 0))   
          )

```




In the original file, there are no **stat.values** but for the sake of the exercise we'll generated some. 

```{r}
## Parameters

beta1 <- 0.5
beta2 <- 2
beta3 <-10

 processedTradeStats  <- processedTradeStats %>%
  mutate(
    epsilon = rnorm(88,50,9),
    Stat..Value =  beta1*CP4 + beta2 * as.numeric(as.factor(EXPORT.REGION))+ 
      beta3 * as.numeric(as.factor(EXPORT.COUNTRY))* Year + epsilon * Year
  )

```

In a classical linear model,
$$
 stat.value  =  \beta_1 * Var1 +  \beta_2 * Var_2 +  \beta_3 * Var_3 + \epsilon
$$

The *stats.value* was generated with the following parameters $\beta_1$ = `r beta2` and  $\beta_2$ = `r beta3`, wich leads to:  

$$
 stat.value  =  `r beta1` * CP4 + `r beta2` * 1_[export region] + `r beta3` * 1_[export country] + \epsilon
$$

# Statistics
Now let's look at our constructed data set and start exploring: 
```{r, eval=FALSE}
head(processedTradeStats)
```
... and let's have some **descriptive statistics**
```{r}
datasummary_skim(processedTradeStats, type = "numeric")
```



## Main Exports over the whole period

### As a table

```{r  }
# Calculate total statistical value by principle export type
totalCommodityValueByPrincipleExportType <- processedTradeStats %>%
  group_by(Principle.Exports) %>%
  summarise(total = sum(Stat..Value)) %>%
  arrange(desc(total))

kable(totalCommodityValueByPrincipleExportType, caption = "Exports from Vanuatu [ 2000-2010].") %>% 
  kable_styling() 

```


 Or as a graphic: 

```{r }
# Plot the exports
totalCommodityValueByPrincipleExportType  %>%
  ggplot(aes(x = reorder(Principle.Exports, total), y = (total/1000000))) +
    geom_bar(stat = "identity", color="black", fill= "lightsteelblue1", position=position_dodge()) +
      labs( x= "Exports categories", y = "Vatu (millions)") +
  ggtitle(label = " Exports from Vanuatu [ 2000-2010] ",
          subtitle = "Vanuatu National Statistical Office")+
  coord_flip() +
  theme_minimal()

```

##  Selection of the Year

Imagine now the same report but for any year $ \in [2000, 2010]$

```{r}
MyYear <- 2003
```

| we have selected the year **`r MyYear`** let's see the report!


### As a table

```{r  }
# Now only for a specific year
NewtotalCommodityValueByPrincipleExportType <- processedTradeStats %>%
  filter(Year == MyYear) %>%
  group_by(Principle.Exports) %>%
  summarise(total = sum(Stat..Value)) %>%
  arrange(desc(total))

kable(NewtotalCommodityValueByPrincipleExportType,
      caption = paste("Exports from Vanuatu, year = ", MyYear)) %>% 
  kable_styling()



```


 Or as a graphic: 

```{r }

# Plot the exports for the specific year
NewtotalCommodityValueByPrincipleExportType  %>%
  ggplot(aes(x = reorder(Principle.Exports, total), y = (total/1000000))) +  
    geom_bar(stat = "identity", color="white", fill= SIAP.color, position=position_dodge()) +
      labs( x= "Exports categories", y = "Vatu (millions)") +
  ggtitle(label = paste("Exports from Vanuatu, year = ", MyYear),
          subtitle = "Vanuatu National Statistical Office")+
  coord_flip() +
  theme_minimal()

```






