---
title: "Trade Data Construction"
author: "Christophe Bontemps"
date: "2024-01-31"
output:
  html_document:
    df_print: paged
    toc: yes
    keep_md: yes
    code_folding: show
    fig_width: 6.5
    fig_height: 4
  pdf_document:
    df_print: kable
    toc: no
    keep_tex: yes
    fig_width: 6.5
    fig_height: 4
  word_document:
    toc: yes
---


```{r setup, include=FALSE}
# Set echo=FALSE as default for code cells
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      dpi = 600)
# My colors:
SIAP.color <- "#0385a8"

```


```{r libraries, include = FALSE}
#### Preparation ####

# Load the required libraries
library(dplyr) # Manipulating data
library(stringr) # common string operations
library(tidyverse) # Manipulating data

# For tables 
library(kableExtra)
library(modelsummary)
library(gt) 
library(gtsummary)
library(tinytable)


 # to get tinytable library(remotes) install_github("vincentarelbundock/tinytable")
```


```{r data, include = FALSE}
# # Note where VNSO code/data is on current computer
# repository <- file.path(dirname(rstudioapi::getSourceEditorContext()$path), "Data")
# setwd(repository) # Required for file.choose() function

# Read in the raw trade data from secure folder of the repository 
tradeStatsFile <- "VBoS_Dummy-Data_16.02.23.csv"
tradeStats <- read.csv(paste0("Data/", tradeStatsFile), header=TRUE, na.strings=c("","NA", "NULL", "null")) #replace blank cells with missing values-NA
```

# Data used 
We use some anaonyized data from Vanuatu National Statistical Office (VNSO) trade data.  and specifically used the file named  *`r tradeStatsFile`*, with  **`r nrow(tradeStats)`** observations on several goods exported.

```{r preparation, include = FALSE}
#### Clean and process the latest month's data ####

# Initial re-formatting of the data

# Remove duplicated rows from the trade statistics data
duplicatedRows <- duplicated(tradeStats) 
tradeStatsNoDup <- tradeStats[duplicatedRows == FALSE, ]

# Convert the statistical value to numeric - note numbers formatted with commas in them and these need removed
tradeStatsNoDup$Stat..Value <- as.numeric(gsub(",", "", tradeStatsNoDup$Stat..Value))

# Convert excel figures to dates (choose)
tradeStatsNoDup$Reg..Date <- as.Date(tradeStatsNoDup$Reg..Date, format = "%d/%m/%Y")

# Create subset for Export and Import commodities 
tradeStatsSubset <- tradeStatsNoDup[tradeStatsNoDup$Type %in% c("EX / 1","EX / 3", "IM / 4", "IM / 7"), ]
tradeStatsCommodities <- tradeStatsSubset[tradeStatsSubset$CP4 %in% c(1000, 4000, 4071, 7100), ]
```


```{r NewVariables, include = FALSE}
#### Extract the Year, Month, and Day as separate columns ####

# Create new columns for dates
tradeStatsCommodities$Year <- format(tradeStatsCommodities$Reg..Date, format= "%Y")
tradeStatsCommodities$Month <- format(tradeStatsCommodities$Reg..Date, format= "%B")
tradeStatsCommodities$Day <- format(tradeStatsCommodities$Reg..Date, format= "%d")

#### Finish Processing ####

# Make copy of latest month's processed data
processedTradeStats <- tradeStatsCommodities
```

### Addition to the original file

The dates and year are missing, so we'll generate the year to have some variabilty
```{r}
 processedTradeStats  <- processedTradeStats %>%
  mutate (Year = as.integer(2000 + round(10*runif(nrow(.)), 0))   
          )

```

In the original file, there are no **stat.values** but for the sake of the exercise we'll generated some. 

```{r}
## Parameters

beta1 <- 0.5
beta2 <- 2
beta3 <-10

 processedTradeStats  <- processedTradeStats %>%
  mutate(
    epsilon = rnorm(88,50,9),
    Stat..Value =  round((beta1*CP4 + beta2 * as.numeric(as.factor(EXPORT.REGION))+ 
      beta3 * as.numeric(as.factor(EXPORT.COUNTRY))* Year + epsilon * Year)/10^3, 2)
  )

```

# Cleaning Variables names

```{r}
# Rename variables to remove dots and replace by "_"
processedTradeStats <-  processedTradeStats %>%
  rename_all(~ gsub("\\.+", "_", .))
```



```{r write, include = FALSE}
# Create csv of last months processed data
write.csv(processedTradeStats, "Data/TradeData.csv")

```
