---
title: "International Merchandise Trade Statistics Highlights"
author: "Your name"
subtitle: '*A simple report*'
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

# Data analysis
We provide here some basic elements about the data set and descriptive statistics on some variables of interest.  

* First we load the data
* second we print a summary

```{r}
TradeData <- read.csv("https://www.unsiap.or.jp/on_line/2024/RAP/Data/TradeData.csv")
```

We loaded the data set and compile the  report on `r Sys.Date()`

## Summary statistics
Here are some summary statistics of the export value (*Stat_Value* variable), all products combined:

```{r }
summary(TradeData$Stat_Value)
```

The data set contains   **`r  nrow(TradeData)`  observations**, and  **`r  ncol(TradeData)`** variables.

Here is a list of all variables available in the data set: 

```{r}
names(TradeData)
```

## Type of exports

The variable *Principle_Export* list all the types of goods exported from Vanuatu. There are **`r length(unique(TradeData$Principle_Exports))`** different goods exported. Let's have the list of all these goods. 

```{r}
unique(TradeData$Principle_Exports)

```






