---
title: "International Merchandise Trade Statistics Highlights"
author: "Christophe Bontemps (SIAP)"
subtitle: '*A simple report*'
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
  pdf_document:
    toc: yes
---



```{r libraries, include = FALSE}
# Load the required libraries

# For manipulating data
library(tidyverse)

# For tables display
library(flextable)

```


```{r setup, include=FALSE}
# Set echo=FALSE as default for code cells
# The code folding for HTLM output is thus deactivated
knitr::opts_chunk$set(echo = FALSE)

```


# Data analysis
We provide here some basic elements about the data set and descriptive statistics on some variables of interest.  

* First we load the data
* second we print a summary

```{r}
TradeData <- read.csv("https://www.unsiap.or.jp/on_line/2024/RAP/Data/TradeData.csv")
```

We loaded the data set and compile the  report on `r Sys.Date()`

## Summary statistics
Here are some summary statistics of the export value (*Stat_Value* variable), all products combined:

```{r }
summary(TradeData$Stat_Value)
```

The data set contains   **`r  nrow(TradeData)`  observations**, and  **`r  ncol(TradeData)`** variables.

Here is a list of all variables available in the data set: 

```{r}
names(TradeData)
```

## Type of exports

The variable *Principle_Export* list all the types of goods exported from Vanuatu. There are **``r length(unique(TradeData$Principle_Exports))`** different goods exported. Let's have the list of all these goods. 

```{r}
unique(TradeData$Principle_Exports)

```

# Main Exports by type of goods

## As a table

```{r  }
# Calculate total value by principle export type
ExportsByType <- TradeData %>%
  group_by(Principle_Exports) %>%
  summarise(total = sum(Stat_Value)) %>%
  arrange(desc(total))

```

```{r}
# Table display
# To reduce the homogenization efforts and the number of functions to be called, 
# it is recommended to define formatting properties once and for all

# set_flextable_defaults(
#   font.size = 10,   table.layout = "autofit")

# Printing the table
flextable(ExportsByType)
```

# As a graphic: 

```{r }
# Plot the exports
ExportsByType  %>%
  ggplot()+
  aes(x = reorder(Principle_Exports, total), y = (total)) +
  geom_bar(stat = "identity", 
             fill= "lightsteelblue1") +
  labs(title ="Exports from Vanuatu [ 2000-2010] ", 
       subtitle = "National Statistical Office", 
       caption  = paste("Date: ",Sys.Date()), 
       x= "Exports categories",
       y = "Vatu") +
  coord_flip() +
  theme_minimal()

```

# Some extensions (not required in the project )

##  Selection of the Year

Imagine now the same report but for any year in [2000, 2010]

```{r}
MyYear <- 2003
```

| we have selected the year **`r MyYear`** let's see the report!


### As a table

```{r  }
# Now only for a specific year
ExportsByType_Year <- TradeData %>%
  filter(Year == MyYear) %>%
  group_by(Principle_Exports) %>%
  summarise(total = sum(Stat_Value)) %>%
  arrange(desc(total))

```


```{r}
flextable(ExportsByType_Year)
```


 Or as a graphic: 

```{r }
# Plot the exports for the specific year
ExportsByType_Year  %>%
  ggplot(aes(x = reorder(Principle_Exports, total), y = (total))) +  
    geom_bar(stat = "identity", color="white", fill= "lightsteelblue1") +
      labs( x= "Exports categories", 
            y = "Vatu", 
            title = paste("Exports from Vanuatu, year = ", MyYear),
            subtitle = "Vanuatu National Statistical Office") +
  coord_flip() +
  theme_minimal()

```





